---
title: "Data Wrangling Challange"
author: "Gabriel Mutua"
date: "August 8, 2018"
output: html_document
---

```{r include=FALSE}
#Load the data provided
challange_1 <- read.csv("~/Rmarkdown/housing_data.csv")
```

```{r include=FALSE}
#load or install all required packages
packages <- c("ggplot2", "rmarkdown", "knitr", "RColorBrewer", "reshape2", "plyr")
if (length(setdiff(packages, rownames(installed.packages()))) > 0) {
  install.packages(setdiff(packages, rownames(installed.packages())))  
}
```


#####**1) How many learners were in the housing trainings project?**

Since the learners ID is unique, I checked the length of the coloumn the learner_id. Although not a good practice we can counter check the results by using dim function, to check the dimension of the whole data frame

```{r}
no_learners = length(challange_1$learner_id)
no_learners
##checking the dimension of the whole data frame.
dim(challange_1)
```

#####**2) Which were the 3 most popular trainings? (Trainings with the most learners)?**

```{r results='asis'}
library(knitr)
popular <- as.data.frame(table(challange_1$program_code))
popular= popular[order(-popular$Freq),]
kable(popular[1:5,], caption = "Top five most popular programs")

```

***Visualizing the top for five popular trainings***

```{r echo=FALSE}
library(ggplot2)
plot = popular[1:5,]
ggplot(data = plot, aes(x = reorder(Var1, -Freq), y = Freq, fill = Var1))+
  geom_bar(stat = "identity")+
  theme(
    axis.title= element_blank(),
    axis.text = element_blank(),
    axis.ticks = element_blank()
  )
```

#####**3) Which learner had the most interactions and how many interactions did they have?**

```{r results='asis'}
library(reshape2)
interactions <- subset(challange_1, select = c("learner_id","variation_code"))
interactions_count <- melt(table(interactions))

library(plyr)
summarized_interaction <- ddply(interactions_count,.(learner_id),summarize,sum=sum(value),number=length(learner_id))
most_interactions = summarized_interaction[order(-summarized_interaction$sum),]
most_interactions = most_interactions[1:5, 1:2]


kable(most_interactions, caption = "Learners with most interactions")

```

***Learners with most interactions***

```{r echo= FALSE}
ggplot(data = most_interactions, aes(x = reorder(as.character(learner_id),-sum), y = sum, fill = as.character(learner_id)))+
  geom_bar(stat = "identity")+
  theme(
    axis.title= element_blank(),
    axis.text = element_blank(),
    axis.ticks = element_blank()
  )
```

From both the table and the plot we can see learner 6484248 had the most interactions

#####**4) How many learners had more than 100 interactions (>100)? **

```{r}
more_than <- subset(summarized_interaction, sum > 100)
length(more_than$learner_id)
```

#####**5a) How many learners had 100 or less interactions (<= 100) **

```{r}
less_than <- subset(summarized_interaction, sum <= 100)
length(less_than$learner_id)
```


#####**5b) derive summary statistics of their interactions for 5a **
```{r}
summary_stats <- summary(less_than$sum)
summary_stats
```

#####**5c) visualize the distribution of their interaction counts of 5a **

```{r warning=FALSE}
qplot(less_than$sum, geom ="histogram")
```

#####**6) Which day of the week had the most interactions and which had the least interactions?**

```{r}
challange_1$day <- weekdays(as.Date(challange_1$created_at))
days = as.data.frame(table(challange_1$day))

days= days[order(-days$Freq),]
kable(days, caption = "Number of interactrions per day")
```

***Interactions per day***

```{r echo=FALSE}
ggplot(data = days, aes(x = reorder(Var1, -Freq), y = Freq, fill = Var1))+
  geom_bar(stat = "identity")+
  theme(
    axis.title= element_blank(),
    axis.text = element_blank(),
    axis.ticks = element_blank()
  )
```

